---
title: "Birthwight"
author: "Esra"
date: "5/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)#this library is used  for most graphs and data functions
library (dplyr)
library(ggrepel)#this library for the labels on the graphs with the lines to the points they represent
```

## Birthwights and gastation for smoker and non somker mothers 

### Research Question
This dataset contains information like gestational age on new born babies and their mothers who smoked and didn’t.  Here we ask,is there a relationship between gestational age and birthweight?

```{r data, include=FALSE} 
# load data
read.csv ("C:\\Users\\ekhojah\\OneDrive\\Desktop\\Birthweight_reduced_R.csv")
# rename data file
df <- read.csv("C:\\Users\\ekhojah\\OneDrive\\Desktop\\Birthweight_reduced_R.csv")
attach(df) 
```

### Data Origins

You can also embed plots, for example:
```{r pressure, echo=FALSE}
head(df)
```

### Data Preparation
For the perpose of this project only the information about Birthweight of the newborn babies, Gestation and their mothers who smoked and didn’t will be adress.
```{r prep, results=TRUE}
df <- df %>% select (-c(headcirumference,mnocig,mppwt,fage,fedyrs,fnocig,fheight,lowbwt,mage35))#exculde culomns we do not need
head(df)#show first few raws of the new dataset
```

### Visulisation 1:


```{r gest,results=TRUE}

plot(Gestation,Birthweight,main="Scatterplot of gestational age and birthweight",xlab="Gestation(weeks)",ylab="Birthweight(lbs)")
abline(lm(Birthweight~Gestation),col='red',lwd=2)
```

## Including Plots

You can also embed plots, for example:

```{r , echo=FALSE}
plot(pressure)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
