---
title: "Birthwight"
author: "Esra"
date: "5/23/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE,fig.show="hold",warning=FALSE,results=TRUE,message=FALSE) 

library(tidyverse)#library for most graphs and data functions
library (dplyr)#library makes steps of execute the program fast and easy
library(ggrepel)#library for the labels on the graphs with the lines 
```

## Infants birthweight and gestational age for smoker and nonsmoker mothers 


### Research Questions

Here we ask,is there a relationship between gestational age and birthweight?
and do smokers have lighter babies?


### Data Origins

This dataset consists of details of 42 babies and their parents at birth.You can find this dataset in [Stastutor](http://www.statstutor.ac.uk/topics/r/statistical-analyses-using-r/).
These data were collected  by the employer's liability insurances in Germany.
As reported in its resear artical by [Karmaus and Wolf ,1995](https://ehp.niehs.nih.gov/doi/abs/10.1289/ehp.951031120).

Variables meaning:
Birthweight= Weight of baby by pound (lbs),
Gestation= Gestation age of baby by (weeks),
smoker= Mother smokes setuation : 1 = smoker , 0 = non-smoker



### Data Preparation

```{r data, include=TRUE, results=FALSE}
# load data
read.csv ("C:\\Users\\ekhojah\\OneDrive\\Desktop\\Birthweight_reduced_R.csv")
# rename data file
df <- read.csv("C:\\Users\\ekhojah\\OneDrive\\Desktop\\Birthweight_reduced_R.csv")
```

To answer the research questions, only the information about birthweight , gestation of the newborn babies and their mothers who smoked and didn’t will be adress.
```{r prep, include=TRUE,results=TRUE}
#exclude all except variables we need
df1 <-df %>% select(-c(id,headcirumference,length,motherage,mnocig,mheight,mppwt,fage,fedyrs,fnocig,fheight,lowbwt,mage35,LowBirthWeight))
#showe first few rews
head(df1)
```

### Visulisation 1:
Here I adress the first question:is there a relationship between gestational age and birthweight?
```{r gest,include=FALSE}
attach(df1)#use the birthweight dataset until further notice using attach.
```

```{r gestation,results=TRUE,warning=FALSE,message=FALSE}
plot(Gestation,Birthweight,main="Scatterplot of gestational age and birthweight",xlab="Gestation(weeks)",ylab="Birthweight(lbs)")
abline(lm(Birthweight~Gestation),col='red',lwd=2)
#save output
png(filename= "birthwight.png")

```

There is a positive relationship, as the gestational age increases, so does birthweight. The closer the scatter is to the line, the stronger the relationship is.


## Summary
This data visualisation showes information of birthwight and gastation age for 42 babies and the effects of thier mothers smoking siuation.
From visulisation 1, we can see there are no large outliers wich could have large impact on the strength of a relationship.  
